---
title: "WildFire Tracker"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: menu
    theme: bootstrap
runtime: shiny
---



States
=======================================================================

Sidebar {.sidebar}
----------------------------------------------------------------------

```{r}
#Define inputs for Drop Down Menu
 selectInput("State", label = h3("Select a State"), 
    choices = list("AL" = 1, "AK" = 2, "AZ" = 3, "AR" = 4, "CA" = 5),
    selected = 20)
```

### Introduction

Row
-----------------------------------------------------------------------

### Fire Distribution

```{r}
#Insert Code for US Fire Distribution Map Here
```


### Fire Size by Census Tract
```{r}
#Insert Code for US Fire Size by Census Tract Map Here
```

Data Visualizations
=======================================================================
Sidebar {.sidebar}
----------------------------------------------------------------------

```{r}
#Define inputs for Drop Down Menu
 selectInput("State", label = h3("Select a State"), 
    choices = list("AL" = 1, "AK" = 2, "AZ" = 3, "AR" = 4, "CA" = 5),
    selected = 20)
```

Row
-----------------------------------------------------------------------
### Data Viz #1
```{r}
#Insert code for data viz #1
```

### Data Viz #2
```{r}
#Insert code for data viz #2
```
### Data Viz #3
```{r}
#Insert code for data viz #3
```
### Data Viz #4
```{r}
#Insert code for data viz #4
```

Row
-----------------------------------------------------------------------
### Data Viz #5
```{r}
#Insert code for data viz #5
```
### Data Viz #6
```{r}
#Insert code for data viz #6
```


About
=======================================================================

**Notes:**

* There are some duplicate rows

* Some rows are missing weather data

* Not all of the columns are documented in the documentation file

* Some of the weather data is completely unrealistic: 

For example, if you sort Temp_pre_30 by lowest to highest, it lists -49°C (-56°F) in Tennessee in November. The lowest recorded temp ever in TN is -32°F.  Also, if you sort highest to lowest and look a few columns over it lists 52°C (125°F) for Wyoming in September.  The highest recorded temperature ever in WY is 115°F. 

* This sample does not include any fires that are class A, which means greater than 0 but less than or equal to 0.25 acres, and there are a significant number of them in the big original data set (almost ⅓ of all fires)

**References:**

* https://stackoverflow.com/questions/36020146/how-can-i-order-the-months-chronologically-in-ggplot2-short-of-writing-the-month

* https://stackoverflow.com/questions/22058393/convert-a-numeric-month-to-a-month-abbreviation/22058435

**Original Dataset**

* https://www.kaggle.com/capcloudcoder/us-wildfire-data-plus-other-attributes?select=Wildfire_att_description.txt




