---
title: "Data Visualization"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(lubridate)
library(plotly)
library(scales)
```

```{r, include = FALSE}
# Load files
wildfires <- read_csv("Abbie-Code/FW_Veg_Rem_Combined.csv")
wildfires$disc_clean_date <- as.Date(wildfires$disc_clean_date, format = "%m/%d/%Y")
```

```{r, include = FALSE}
# Create initial weather data frame
weather_data <- wildfires %>% 
  filter(weather_file != "File Not Found") %>% 
  mutate(temp_f = ((9/5)*Temp_cont + 32)) %>% 
  mutate(wind_mph = 2.237*Wind_cont) %>% 
  mutate(precip_in = Prec_cont/25.4)

# Filter rows where all the weather-related values are 0
filtered_weather_data <- weather_data %>% 
  filter(Temp_pre_30 != 0.00000000 | Temp_pre_15 != 0.00000000 | Temp_pre_7 != 0.00000000 | Temp_cont != 0.00000000 | Wind_pre_30 != 0.00000000 | Wind_pre_15 != 0.00000000 | Wind_pre_7 != 0.00000000 | Wind_cont != 0.00000000 | Hum_pre_30 != 0.00000000 | Hum_pre_15 != 0.00000000 | Hum_pre_7 != 0.00000000 | Hum_cont != 0.00000000 | Prec_pre_30 != 0.00000000 | Prec_pre_15 != 0.00000000 | Prec_pre_7 != 0.00000000 | Prec_cont != 0.00000000) 

# Filter only the wildfires that are bigger than 1000 acres
big_wildfires <- wildfires %>% 
  filter(fire_size_class == "F" | fire_size_class == "G")

# Filter only the wildfires that are bigger than 1000 acres from the weather data
big_wildfires_weather <- filtered_weather_data %>% 
  filter(fire_size_class == "F" | fire_size_class == "G")
```

$~$
$~$
$~$

```{r, echo = FALSE}
# Average Size of Fire Each Year
average_size_year <- wildfires %>% 
  group_by(year(disc_clean_date)) %>% 
  summarize(mean_size = mean(fire_size))

average_size_plot <- ggplot(average_size_year, aes(x = `year(disc_clean_date)`, y = mean_size)) +
  geom_line() +
  geom_smooth(method = "lm", se = F) +
  labs(title = "Average Size of Fire Each Year, 1992-2015", x = "Year", y = "Average Size of Fire (in Acres)")
ggplotly(average_size_plot)
```

  > This graph shows that between 1992 and 2015, there was a significant upward trend in the average size of wildfires in the United States.
  
$~$
$~$
$~$


```{r, echo = FALSE}
# Cause vs. Average Size of Fire
mean_size_cause <- wildfires %>% 
  group_by(stat_cause_descr) %>% 
  summarize(average_size = mean(fire_size))

mean_size_cause_plot <- ggplot(mean_size_cause, aes(x = reorder(stat_cause_descr, -average_size), y = average_size)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Average Size of Fire, by Supposed Cause, 1992-2015", x = "Supposed Cause of Fire", y = "Average Size of Fire (in Acres)")
ggplotly(mean_size_cause_plot)
```

  > This graph suggests that lightning strikes caused considerably larger fires between 1992 and 2015.
  
$~$
$~$
$~$


```{r, echo = FALSE, fig.show = "hold", out.width = "50%"}
# Number of Fires Each Year
discovery_year <- wildfires %>% 
  group_by(year(disc_clean_date)) %>% 
  count(year(disc_clean_date))

num_year_plot <- ggplot(discovery_year, aes(x = `year(disc_clean_date)`, y = n)) +
  geom_line() +
  labs(title = "Total Number of Fires Recorded Each Year, 1992-2015", x = "Year", y = "Number of Fires") +
  geom_smooth(method = "lm", se = F)
num_year_plot


# Number of Large Fires Each Year
big_discovery_year <- big_wildfires %>% 
  group_by(year(disc_clean_date)) %>% 
  count(year(disc_clean_date))

big_num_year_plot <- ggplot(big_discovery_year, aes(x = `year(disc_clean_date)`, y = n)) +
  geom_line() +
  labs(title = "Number of Fires >1000 Acres Recorded Each Year, 1992-2015", x = "Year", y = "Number of Fires") +
  geom_smooth(method = "lm", se = F)
big_num_year_plot
```

  > (Note the different scales on the y-axes) This comparison indicates that even though there was not a large increase in the total number of fires between 1992 and 2015, there was a more significant upward trend in the number of large fires greater than 1000 acres during that same period.
  
$~$
$~$
$~$


```{r, echo = FALSE}
# Acreage Burned Per Year
acreage_year <- wildfires %>% 
  group_by(year(disc_clean_date)) %>% 
  summarize(fire_size = sum(fire_size))

acreage_year_plot <- ggplot(acreage_year, aes(x = `year(disc_clean_date)`, y = fire_size)) +
  geom_line() +
  geom_smooth(method = "lm", se = F) +
  scale_y_continuous(label = number_format(big.mark = "")) +
  labs(title = "Acreage Burned Per Year, 1992-2015", x = "Year", y = "Number of Acres Burned") 
ggplotly(acreage_year_plot)
```

  > This graph shows that the total number of acres burned substantially increased during the period from 1992 to 2015. 
  
$~$
$~$
$~$

```{r, echo = FALSE, message = FALSE, warning = FALSE}
year_month_breakdown <- wildfires %>% 
  group_by(month(disc_clean_date), year(disc_clean_date)) %>% 
  summarize(total_fire_size = sum(fire_size))

year_month_breakdown <- year_month_breakdown %>% 
  mutate(date = make_date(year = `year(disc_clean_date)`, month = `month(disc_clean_date)`))

layout_ggplotly <- function(gg, x = -0.02, y = -0.1){
  # The 1 and 2 goes into the list that contains the options for the x and y axis labels respectively
  gg[['x']][['layout']][['annotations']][[1]][['y']] <- x
  gg[['x']][['layout']][['annotations']][[2]][['x']] <- y
  gg
}

year_month_size_plot <- ggplot(year_month_breakdown, aes(x = date, y = total_fire_size)) +
  geom_col() +
  scale_x_discrete(limits = month.abb) +
  labs(title = "Acreage Burned Each Month, 1992-2015", x = "Month", y = "Total Acreage Burned", fill = "Month") +
  scale_y_continuous(labels = number_format(big.mark = "")) +
  facet_wrap(~`year(disc_clean_date)`, scales = "free_x", nrow = 2)
ggplotly(year_month_size_plot) %>% 
  layout_ggplotly()
```

  > This graph shows the seasonal trend of wildfires, along with the upward trend in how much more area is burned by wildfires by the end of this approximately 20-year period.
  
$~$
$~$
$~$


